---
title: "EDA"
author: "Sri Santhosh Hari"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(magrittr)
library(tseries)
library(forecast)

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r, echo=F}
# Load data and add Mon column in date format
train_data <- read_csv("Data/train.csv")
train_data <- na.omit(train_data)
train_data$Mon <- seq.Date(as.Date("1987/1/1"), as.Date("2010/12/1"), by = "month")
```

We first visualize how the variable of importance, `Bankruptcy Rate`, has changed over the years.

```{r, echo=F, fig.height=4, fig.cap="\\label{fig:BRTrend}Monthly Canadian Bankruptcy Rates (1987 - 2010)"}
train_data %>% 
  ggplot(aes(x=Mon))+
  geom_line(aes(y=scale(Bankruptcy_Rate)))+
  labs(x='Time', y='Bankruptcy Rate', title='')
```

It can be observed from Figure \ref{fig:BRTrend} that the bankruptcy rates have generally increased since 1987 with significant increase in later half of 2000's, which can be attributed to the global financial crisis/recession.

```{r, echo=F, fig.height=4, fig.cap="\\label{fig:BRMonth}Bankruptcy Rates across Months"}
train_data %>% 
  ggplot(aes(x=factor(lubridate::month(train_data$Mon))))+
  geom_boxplot(aes(y=Bankruptcy_Rate,alpha=0.5),show.legend=F)+
  scale_x_discrete(labels=substr(month.name,1,3))+
  labs(x="Month", y="Monthly Bankruptcy Rate", title="")
```

Looking at the monthly changes in Bankruptcy Rates in Figure \ref{fig:BRMonth}, one can infer that March, the last month of financial year, tends to have higher bankruptcy rates compared to other months.

```{r, echo=F, fig.height=4, fig.cap="\\label{fig:Trend}Trends observed in variables across years"}
train_data %>% 
  ggplot(aes(x=Mon))+
  geom_line(aes(y=scale(Bankruptcy_Rate), color='Bankruptcy Rate'))+
  geom_line(aes(y=scale(Unemployment_Rate), color='Unemployment Rate'))+
  geom_line(aes(y=scale(Population), color='Population'))+
  geom_line(aes(y=scale(House_Price_Index), color='House Price Index'))+
  labs(x='Time', y='Scaled Variable', color='')
```

We then visualize the changes observed in external variables (House Price Index, Population, Unemployment Rate) to asses if multivariate approach is suitable as opposed to univariate modeling. Figure \ref{fig:Trend} shows the trend in each of the external variable (scaled) along with Bankruptcy rate. It can be seen that HPI has a clear relationship with Bankruptcy Rate, as the trends observed in both the variables appear to be similar, with some time-difference. Population appears to linearly increase over the years but without valleys, may not be an ideal variable to include in the model. We will evaluate models with population included and excluded.

```{r, echo=F, fig.height=4, fig.cap="\\label{fig:BRHPI}Bankruptcy Rates and House Price Index changes across years"}
train_data %>% 
  ggplot(aes(x=Mon))+
  geom_line(aes(y=scale(Bankruptcy_Rate), color='Bankruptcy Rate'))+
  geom_line(aes(y=scale(House_Price_Index), color='House Price Index'))+
  geom_line(aes(y=scale(dplyr::lag(train_data$House_Price_Index,18)),
                color='Lagged HPI'))+
  labs(x='Time', y='Scaled Variable', color='')
```

Figure \ref{fig:BRHPI} helps us visualize the the relationship of HPI and 18-month-lagged verison of HPI with Bankruptcy Rate.

```{r, echo=F, fig.cap="\\label{fig:scatter}Bankruptcy vs External Variables"}
p1 <- train_data %>% ggplot() +
  geom_point(aes(x=Bankruptcy_Rate,y=House_Price_Index,alpha=0.5), show.legend = F)+
  labs(x='Bankruptcy Rate',y='House Price Index')
p2 <- train_data %>% ggplot() +
  geom_point(aes(x=Bankruptcy_Rate,y=dplyr::lag(train_data$House_Price_Index,18),
                 alpha=0.5), show.legend = F)+
  labs(x='Bankruptcy Rate',y='House Price Index(10 mon lagged)')
p3 <- train_data %>% ggplot() +
  geom_point(aes(x=Bankruptcy_Rate,y=Population, alpha=0.5), show.legend = F)+
  labs(x='Bankruptcy Rate',y='Population')
p4 <- train_data %>% ggplot() +
  geom_point(aes(x=Bankruptcy_Rate,y=Unemployment_Rate, alpha=0.5), show.legend = F)+
  labs(x='Bankruptcy Rate',y='Unemployment Rate')
multiplot(p1,p3,p2,p4, cols=2)
```

Figure \ref{fig:scatter} visually captures the relationship between the Bankruptcy Rate and other external variables available.

```{r, echo=F}
# Correlation Check
cor_tab <- as.data.frame(round(cor(train_data[2:5])*100,2))
cor_tab %<>% 
  select(Bankruptcy_Rate) %>%
  filter(Bankruptcy_Rate != 100)

#Lagged HPI
train_mod <- train_data[2:5]
train_mod$HPI_lag <- dplyr::lag(train_mod$House_Price_Index,18)
lagged_cor <- round(100*cor(train_mod$Bankruptcy_Rate[c(19:288)], train_mod$HPI_lag[c(19:288)]),2)

cor_tab <- rbind(cor_tab, setNames(as.data.frame(lagged_cor),c("Bankruptcy_Rate")))

colnames(cor_tab) <- c("Bankruptcy Rate")
rownames(cor_tab) <- c("Unemployment Rate","Population","House Price Index", "Lagged House Price Index")

knitr::kable(cor_tab, caption = "\\label{tab:corr}Correlation (in %) of Bankruptcy Rate with External Variables")
```

